---
title: "markdown_onetwo"
author: "Victor Møller"
date: "20 okt 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#libraries
library(pacman)
p_load(tidyverse, lme4, corrplot, RColorBrewer, lmerTest, effects, MuMIn, cowplot) 

#loading data
data <- read.csv("language_dev_asd_clean_2.csv") #data from assignment 1 

#right format 
str(data) 

#conversion
data$SUBJ <- as.factor(data$SUBJ) 
data$VISIT <- as.factor(data$VISIT)

#well matched? 
participants <- data %>%
  filter(VISIT == 6)

summary(participants$Diagnosis) #29 ASD & 32 TD 

#creating subsets for age 
data_ASD <- data %>%
  filter(!is.na(Age) & VISIT == 1 & Diagnosis == "ASD") 

data_TD <- data %>%
  filter(!is.na(Age) & VISIT == 1 & Diagnosis == "TD") 

#z-score for ASD
data_ASD <- data_ASD %>% 
  mutate(ageZ = scale(Age))

data_TD <- data_TD %>%
  mutate(ageZ = scale(Age))

#summary ASD
summary(data_ASD$ageZ) #bigger z-scores, some very old and young 
which(data_ASD$ageZ > 2) # 
which(data_ASD$ageZ < -2) #8

#summary TD
summary(data_TD$ageZ) #somewhat smaller z-scores 
which(data_TD$ageZ > 2) #2, 29 
which(data_TD$ageZ < -2) #none

#comparing means 
difference <- mean(data_TD$Age) - mean(data_ASD$Age)
difference #huge difference of 12.6 (months older in general)

#comparing means with Z
z_diff <- mean(data_TD$ageZ) - mean(data_ASD$ageZ)
z_diff #very very small. 

#histograms
hist(data_ASD$ageZ) #we have one huge outlier (younger)
hist(data_TD$ageZ) #A couple of outliers (older)

hist(data_ASD$Age) 
hist(data_TD$Age) #don't get quite as old - matching production. 
summary(data_ASD$Age) #18.77 - 42

#Gender
summary(data_ASD$Gender) #Female = 146, Male = 26
summary(data_TD$Gender) #Female = 155, Male = 35

#Ethnicity
summary(data_ASD$Ethnicity) #ASD mainly white & some scattered. 
summary(data_TD$Ethnicity) #TD only white & asian 

#random intercepts
data$Kid = factor(data$SUBJ) #from lecture 

#plotting w. colors
ggplot(data, aes(as.numeric(VISIT), CHI_MLU, color = Diagnosis)) +
  geom_point() + 
  geom_smooth(method = "lm")

#boxplot 
ggplot(data, aes(Diagnosis, CHI_MLU))+
  geom_boxplot() #outliers in ASD. 

#beeswarm plot
p_load(beeswarm)

boxplot(CHI_MLU ~ Diagnosis, data = data, 
  outline = FALSE,    
  main = 'boxplot + beeswarm')
beeswarm(CHI_MLU ~ Diagnosis, data = data, 
  col = 4, pch = 16, add = TRUE)

#violin plot 
ggplot(data, aes(Diagnosis, CHI_MLU)) +
  geom_violin(aes(fill = CHI_MLU)) +
  geom_boxplot(width = 0.2)

#linear mixed effects model 
data$VISIT <- as.integer(data$VISIT) #crucial to run the model 

model_data <- data %>%
  filter(!is.na(CHI_MLU) & !is.na(VISIT) & !is.na(Diagnosis))

#different potential models 
library(MuMIn)

mixed_model <- lmer(CHI_MLU ~ VISIT * Diagnosis + (1+VISIT|SUBJ), data = model_data, REML=FALSE)
summary(mixed_model) #VISIT very significant, DiagnosisTD not significant. 
r.squaredGLMM(mixed_model) #R2m = fixed, R2c = fixed + random. 

#Other models 
model_simple <- lmer(CHI_MLU ~ VISIT + (1+VISIT|SUBJ), model_data, REML=FALSE)
model_diagnosis <- lmer(CHI_MLU ~ Diagnosis + (1+VISIT|SUBJ), data = model_data, REML=FALSE)
model_noint <- lmer(CHI_MLU ~ VISIT + Diagnosis + (1+VISIT|SUBJ), data = model_data, REML=FALSE)

#pairwise anova tests 
anova(model_simple, mixed_model) #mixed model not sign. better
anova(model_diagnosis, mixed_model) #mixed model sign. better
anova(model_noint, mixed_model) #model_simple sign. better

#r-squared for the two best models (model_simple, mixed_model)
r.squaredGLMM(mixed_model) #mixed model has highest marginal r-squared. 
r.squaredGLMM(model_noint) 

#visualizing quadratic 
plot2 <- ggplot(data, aes(as.numeric(VISIT), CHI_MLU, color = Diagnosis)) +
  geom_point() + 
  stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  labs(x = "visit", y = "", legend = "", title = "quadratic")

#visualizing Cubic
plot3 <- ggplot(data, aes(as.numeric(VISIT), CHI_MLU, color = Diagnosis)) +
  geom_point() + 
  stat_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  labs(x = "visit", y = "", title = "cubic")

#visualizing linear
plot1 <- ggplot(data, aes(as.numeric(VISIT), CHI_MLU, color = Diagnosis)) +
  geom_point() + 
  stat_smooth(method = "lm") +
  labs(x = "visit", y = "child mean length of utterance (MLU)", title = "linear")

#spaghetti plot
ggplot(data, aes(VISIT, CHI_MLU, slope = SUBJ, color = Diagnosis)) +
  geom_smooth() #disgusting. 

#different models 
m_linear <- lmer(CHI_MLU ~ VISIT*Diagnosis + (1+VISIT|SUBJ), data = model_data, REML=FALSE)

m_quadratic <- lmer(CHI_MLU ~ (VISIT + I(VISIT^2))*Diagnosis + (1+VISIT + I(VISIT^2)|SUBJ), data = model_data, REML=FALSE)

m_cubic <- lmer(CHI_MLU ~ (VISIT+ I(VISIT^2)+ I(VISIT^3))*Diagnosis + (1+VISIT + I(VISIT^2) + I(VISIT^3)|SUBJ), data = model_data, REML=FALSE)

#evaluating them 
library(lmerTest)
summary(m_cubic) #cubic seems to perform poorly (i.e., have non-significant effects)
summary(m_quadratic) #quadratic seems to perform well (i.e., have significant effects)
summary(m_linear)

#comparing them 
anova(m_linear, m_quadratic) #quadratic best
anova(m_quadratic, m_cubic) #qubic best
anova(m_linear, m_cubic) #qubic best 

#comparison of r-squared 
r.squaredGLMM(m_quadratic) # R2m = .23
r.squaredGLMM(m_cubic) # R2m = .22
r.squaredGLMM(m_linear) # R2m = .22

#effects package plotting (quadratic)
library(effects)
ef <- effect("VISIT:Diagnosis", m_quadratic, xlevels = 6)
summary(ef) #visit 3 

eff_plot <- as.data.frame(ef)
eff_plot

ggplot(eff_plot, aes(VISIT, fit, color=Diagnosis)) + geom_point() + geom_errorbar(aes(ymin=fit-se, ymax=fit+se), width=0.4) + theme_bw(base_size=12)  

#this only gives us 5 levels
plot(predictorEffects(m_quadratic))

#visualizing lineaer
ggplot(data, aes(as.numeric(VISIT), MOT_MLU, color = Diagnosis)) +
  geom_point() + 
  stat_smooth(method = "lm") +
  labs(x = "visit", y = "mother mean length of utterance (MLU)")

#visualizing quadratic 
ggplot(data, aes(as.numeric(VISIT), MOT_MLU, color = Diagnosis)) +
  geom_point() + 
  stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  labs(x = "visit", y = "mother mean length of utterance (MLU)")

#visualizing Cubic
ggplot(data, aes(as.numeric(VISIT), MOT_MLU, color = Diagnosis)) +
  geom_point() + 
  stat_smooth(method = "lm", formula = y ~ poly(x, 3))

#different models 
mot_linear <- lmer(MOT_MLU ~ VISIT*Diagnosis + (1+VISIT|SUBJ), data = model_data, REML=FALSE)

mot_quadratic <- lmer(MOT_MLU ~ (VISIT + I(VISIT^2))*Diagnosis + (1+VISIT + I(VISIT^2)|SUBJ), data = model_data, REML=FALSE)

mot_cubic <- lmer(MOT_MLU ~ (VISIT+ I(VISIT^2)+ I(VISIT^3))*Diagnosis + (1+VISIT + I(VISIT^2) + I(VISIT^3)|SUBJ), data = model_data, REML=FALSE)

#summaries
summary(mot_cubic) #needs interpretation
summary(mot_quadratic) #needs interpretation 
summary(mot_linear)

#comparison
anova(mot_linear, mot_quadratic, mot_cubic)

#comparison pairwie 
anova(mot_quadratic, mot_cubic) #cubic not significantly better
anova(mot_linear, mot_quadratic) #quadratic significantly better 

#r-squared
r.squaredGLMM(mot_quadratic) # R2m = .24
r.squaredGLMM(mot_cubic) # R2m = .23
r.squaredGLMM(mot_linear) # R2m = .23 

#initial correlation
cor_data = select(model_data, VISIT, Age, ADOS_1, nonVerbalIQ_1, verbalIQ_1, MOT_MLU, types_MOT, tokens_MOT, types_CHI, tokens_CHI, CHI_MLU) %>%
  filter(!is.na(Age))

corr = round(cor(cor_data,method = "spearman"),2)

#stepwise w. lmer for quadratic  
big_model <- lmer(CHI_MLU ~ (VISIT + I(VISIT^2))*Diagnosis + (1+VISIT + I(VISIT^2)|SUBJ) + Gender + Ethnicity + ADOS_1 + nonVerbalIQ_1 + verbalIQ_1, data = model_data, REML=FALSE)
summary(big_model)

big_model <- lmer(CHI_MLU ~ (VISIT + I(VISIT^2))*Diagnosis + (1+VISIT + I(VISIT^2)|SUBJ) + verbalIQ_1, data = model_data, REML=FALSE)

(step_res <- step(big_model))
final <- get_model(step_res)
anova(final) 
summary(final) #works, good t-values 

#compared to the quadratic model earlier 
anova(m_quadratic, big_model) 
anova(big_model) 

#r-squared
r.squaredGLMM(big_model) #R2m .6
```

Hypothesis 1 

Plotting 

```{r cars, echo = FALSE, messages = FALSE}
#plotting w. colors
ggplot(data, aes(as.numeric(VISIT), CHI_MLU, color = Diagnosis)) +
  geom_point() + 
  geom_smooth()


```

Initial model - argue for the fact that we chose to include what we did in fact include. 

```{r, echo = FALSE, messages = FALSE}
#mixed model 
summary(mixed_model) #VISIT very significant, DiagnosisTD not significant. 
r.squaredGLMM(mixed_model) #R2m = fixed, R2c = fixed + random. 
```

testing the model with interaction (mixed_model) against the one without (model_noint)

```{r, echo = FALSE, messages = FALSE}
#comparison of mixed model with the same without interaction 
anova(model_noint, mixed_model)

```

plotting models with varrying higher order blabla

```{r, echo=FALSE, messages = FALSE}
#growth curves plot
plot_grid(plot1, plot2, plot3, labels = "AUTO")

```

Seems like the data fits better with quadratic / cubic than linear

```{r, echo=FALSE, messages = FALSE}
#growth curves test 
anova(m_quadratic, m_cubic, m_linear)

```

quadratic chosen because lowest BIC and interpretability 

```{r, echo=FALSE, messages = FALSE}
#r-squared of the chosen model
r.squaredGLMM(m_cubic) 
```

other r-squared values at github. 

```{r, echo=FALSE, messages = FALSE}

```

[report results]

-------------------------------------------------------------------------------------------------

Hypothesis 2

```{r, echo=FALSE, messages = FALSE}
#comparison of models for mommy (beware of BIC)
anova(mot_linear, mot_quadratic, mot_cubic)
```

```{r, echo=FALSE, messages = FALSE}
#r-squared for best model - perhaps replace with linear. 
r.squaredGLMM(mot_quadratic)
```

[report results]

exercise 4
output of the best model (explain how we got here baby)

```{r, echo=FALSE, messages = FALSE}
anova(final) 
```

r-squared for the best model 

```{r}
r.squaredGLMM(final)
```

